Structuring College Access: The Market Segment Model and College Board Geomarkets

### 2025-05-13 text/back_end_KS.qmd
dallas revisions

@fig-dallas-rq2-race presents the racial/ethnic composition of purchased student profiles within Geomarkets in the Dallas-Forthworth area. At the lowest PSAT score thresholds (1070-1180), notable tiers emerge based on the racial/ethnic composition of prospects between Geomarkets. For instance, the central suburban Geomarkets of Dallas County (36% White, 17% Asian, 15% Black, 28% Hispanic) and Irving, Arlington, and Grand Prairie (34% White, 13% Asian, 15% Black, 34% Hispanic) have the greatest proportions of non-White prospects, which is largely attributed to their greater shares of Asian and Black prospects. While the urban central Geomarkets of City of Dallas and City of Fortworth are the most racially/ethnically diverse in the larger metropolitan area (see @fig-dallas-rq1-race), their prospect pools in @fig-dallas-rq2-race includes larger proportion of Hispanic prospects but lower proportions of Asian and Black students comparatively. The Collin & Rockwall Counties (57% White, 7% Black, 17% Hispanic, 14% Asian) and West of Dallas/Forth Worth Metroplex Geomarkets (61% White, 7% Black, 19% Hispanic, 7% Asian) both yield the largest and predominantly White prospect pools.

@@

Prospect pools within Geomarkets from the middle range PSAT score (1190-1260) to the high PSAT score range (1270-1520) become disproportionately more Asian, while the representation of Hispanic and Black prospects decline substantially.

@fig-dallas-rq2-midSAT-race presents the race/ethnicity of all prospects targeted in the middle PSAT score range across Geomarket contributions. @fig-dallas-rq2-midSAT-race shows nearly all Geomarkets contributed to the pool of White prospects at rates nearly proportional to their overall contributions to all included prospects, with the exception of the West of Dallas/Fort Worth Metroplex Geomarket contributing a disproportionately larger share (29% versus 23% for all prospects). Larger disparities across Geomarket contributions were evident for Asian, Black, and Hispanic students. The Dallas County (20% versus 13% for all prospects) and Colin and Rockwall Counties (47% versus 37% for all prospects) Geomarkets contributed a disproportionate larger share of Asian students, whereas nearly all other Geomarkets contributed disproportionately smaller shares.

>>>>

@fig-dallas-rq2-race presents the racial/ethnic composition of purchased student profiles within Geomarkets in the Dallas-Forthworth area, revealing a comparatively more distributed pattern of racial/ethnic diversity across Geomarkets unlike Chicago's concentrated diversity within a single Geomarket. This distribution is most notable at the lowest PSAT score threshold. For instance, at the 1070-1180 PSAT score range, the central suburban Geomarkets of Dallas County (36% White, 17% Asian, 15% Black, 28% Hispanic) and Irving, Arlington, and Grand Prairie (34% White, 13% Asian, 15% Black, 34% Hispanic) have the greatest proportions of non-White prospects, which is largely attributed to their greater shares of Asian and Black prospects. While the urban central Geomarkets of City of Dallas and City of Fortworth are the most racially/ethnically diverse in the larger metropolitan area (see @fig-dallas-rq1-race), their prospect pools in @fig-dallas-rq2-race includes larger proportions of Hispanic prospects but lower proportions of Asian and Black students. The Collin & Rockwall Counties (57% White, 7% Black, 17% Hispanic, 14% Asian) and West of Dallas/Forth Worth Metroplex Geomarkets (61% White, 7% Black, 19% Hispanic, 7% Asian) both yield the largest but predominantly White prospect pools.

@@

Prospect pools within Geomarkets from the middle range PSAT score (1190-1260) to the high PSAT score range (1270-1520) become disproportionately more Asian, while the representation of Hispanic and Black prospects decline substantially. Such shifts also shift distribution patterns of racial/ethnic diversity across Geomarkets, making City of Dallas and City of Fort Worth critical Geomarkets for including Black and Hispanic students from the resulting prospect pools. \[KS LEFT OFF HERE}

@fig-dallas-rq2-midSAT-race presents the race/ethnicity of all prospects targeted in the middle PSAT score range across Geomarket contributions. Disparities in Geomarket contributions were evident but differed for Asian, Black, and Hispanic students. For instance, The Dallas County (20% versus 13% for all prospects) and Colin and Rockwall Counties (47% versus 37% for all prospects) Geomarkets contributed a disproportionate larger share of Asian students, whereas nearly all other Geomarkets contributed disproportionately smaller shares.

<!--# @fig-dallas-rq2-midSAT-race shows nearly all Geomarkets contributed to the pool of White prospects at rates nearly proportional to their overall contributions to all included prospects, with the exception of the West of Dallas/Fort Worth Metroplex Geomarket contributing a disproportionately larger share (29% versus 23% for all prospects).  -->
